Learning Excursion Sets of Vector-valued Gaussian Random Fields for Autonomous Ocean Sampling

### 2021-01-08 ans_to_reviewers/ans-to-2nd-rev.tex
Updates from Overleaf

@@

We are grateful to the editors and reviewers for their comments and
suggestions. They have all been considered in this revised
version of our manuscript.\\

@@

\begin{quote}
"The sampling capabilities are restricted by the sensors onboard
the AUV, the computational load and navigation resources. The sampling
of temperature and salinity variables can be at 1Hz (higher than the
grid resolution), but, due to computational and navigation resources
available on the AUV, this simplification of data assimilation at the
waypoints is necessary."   
\end{quote}

@@

\vspace{5mm}
\reply: We agree that the remarks of Ehrenberg are very useful! (It should be curriculum for graduate students.) Regarding the structure, we seriously considered your suggestions about the background in supplementary material, but in the end thought it was too much of a reduction to go through with. It would be very difficult to grasp the applied contribution of our paper if this background part was in the supplemental material, and only some readers take the extra effort of reading supplemental material. Based on the report of the reviewer, we did however go through the background and cut a few unnecessary and lengthy formulations in the revised version.

@@

\vspace{5mm}
\reply: We agree on all our points here, except Fig.8. Line type will be tricky for Fig. 8, when lines are very close together especially in Fig.8b, and Fig.8c. So we ask to consider that this figure is kept in color. Fig. 4, 5, and 6 have been made with a black-and-white compatible color-map, so they can be converted to black-and-white.

@@

\vspace{5mm}
\reply: We have elaborated on the specification of nugget in the revised version. The sensor-provided standard deviations of the measurement tools are in the fourth decimal points, and in related data-sets we have not seen any nugget effect when measuring temperature and salinity in water samples. We believe that one challenge with the initial sampling line is that it was rather short and that it did no samples at very short distance. We made an empirical variogram from the actual survey data, which are shown below in Figure \ref{empVar}, and these curves support our statements of a smaller nugget.
\begin{figure}[!b] 
\centering
\includegraphics[width=0.99\textwidth]{Figures/field-trials/empVAR.png}
\caption{Empirial variogram from the two runs. }
\label{empVar}
\end{figure}

One reason to keep a significant nugget effect here is to account for the dynamic water masses that are not captured in our model, but we expect this to be very small in our limited survey time. Then again, between the two surveys, there is a change in the survey 1 and 2 paths caused by slightly different measurements at the two times of surveying. Another reason for including a nugget is the uncertainty in the AUV location which means it samples might not be allocated to its true position.

We further checked the sensitivity of the nugget in the sequential algorithm. This was done by running synthetic cases of sampling paths and data up to a certain stage, but then updating the model with different nugget effects in the Gaussian random field model. Figure X shows sampling paths and the radar plots for the AUV direction choice at one steps (top vs bottom) for two different nuggets. Straighter lines when the nugget is large.

In the revised manuscript we wrote another sentence on Fig 11: 'xx'

@@

\vspace{5mm}
\reply: We have improved our

>>>>

\usepackage{float}
\usepackage{cleveref}

@@

We are grateful to the editors and reviewers for their comments and suggestions in this 2$^{nd}$ review. They have all been considered in this revised version of our manuscript. The changes in the manuscript are reflected in the points shown below.\\

@@

\begin{quote}
"The sampling capabilities are restricted by the sensors onboard the AUV, the computational load, and navigation resources. The sampling of temperature and salinity variables can be much higher than the grid resolution, but, due to computational and navigation resources available on the AUV, assimilation into a discretized fixed grid is necessary.`` 
\end{quote}

@@

\vspace{5mm}
\reply: We agree that the remarks of Ehrenberg are very useful! (It should be curriculum for graduate students.) Regarding the structure, we seriously considered your suggestions about the background in supplementary material, but in the end thought it was too much of a reduction to go through with. It would be very difficult to grasp the applied contribution of our paper if the background parts were in the supplemental material, and only some readers take the extra effort of reading that. Based on the report of the reviewer, we did however go through the background and cut a few unnecessary and lengthy formulations in the revised version. {\bf{to go through for typos, redundant text, etc}}

@@

\vspace{5mm}
\reply: We agree on all our points here, except Fig.8. Line type will be tricky for Fig. 8, when lines are very close together especially in Fig.8b, and Fig.8c. So we ask to consider that this figure is kept in color. Fig. 4, 5, and 6 have been made with a black-and-white compatible color-map, so they can be converted to black-and-white. {\bf{to check all}}

@@

\vspace{5mm}
\reply: We have elaborated on the specification of nugget in the revised version. The sensor-provided standard deviations of the measurement tools are in the fourth decimal points, and in related data-sets we have not seen much of a nugget effect for temperature and salinity in water samples close in space and time. The initial sampling transect had no samples at very short distance, and it is hence not possible to see the nugget effect in the empirical variogram. We made an empirical variogram from the actual survey data, where we have some samples close in space, see Figure \ref{empVar}. These curves support our statements of a smaller nugget.
\begin{figure}[H] 
\centering
\includegraphics[width=0.6\textwidth]{Figures/field-trials/empVAR.png}
\caption{Empirical variogram of the scaled variables from the two runs. }
\label{empVar}
\end{figure}

One reason to keep a significant nugget effect here is to account for the dynamic water masses that are not captured in our model, but we expect this to be very small given our limited survey time, as was the case in \cite{fossuminformation}. Then again, between the two surveys, there is a change in the survey 1 and 2 paths caused by slightly different measurements at the two times of surveying. There could also be some uncertainty in the AUV location leading to samples not allocated to its true position.

We further checked the sensitivity of the nugget specification in the sequential sampling algorithm. This was done by running synthetic cases of sampling paths and data up to a certain stage for different measurement noise variance levels. At the end, the posterior excursion probability and the chosen sampling path are compared, as show in Figure \ref{fig:nugget_comparison}. We see moderate changes in the path patterns 
and overall, paths tend to be smoother for large nugget effect.
Indeed, with small nugget, the AUV reacts more to the observed temperature and salinity levels, and this leads to more small turns in its path. 
We note that though the sampling paths can be different, the posterior excursion probabilities are similar and will thus lead to similar estimates for the excursion set. We also remark that the differences in Figure \ref{fig:nugget_comparison} would get smaller if we would let the sampling continue (here stopped for presentation purposes).

\begin{figure}[H]
    \centering
    \includegraphics[scale=0.52]{ans_to_reviewers/merged_paths_2.png}
    \caption{Comparison of sampling paths and posterior excursion probability for different values of the measurement noise standard deviation (from left to right: $0.1$, $0.25$ and $0.5$). The sampling has been stopped after 200 steps.}
    \label{fig:nugget_comparison}
\end{figure}

\newpage

In the revised manuscript we wrote another sentence on Fig 11: 'The AUV paths in both runs are relatively winding in the sense that the vehicle reacts much to the data. In our experience from simulated runs, the paths tend to be smoother when we use a larger nugget effect. Still, the estimated nugget from the actual trial data is about $0.1$ which is close to the specified level.'

@@

\vspace{5mm}
\reply: We have improved our {\bf{to do}}

### 2021-01-08 main_web.tex
Updates from Overleaf

The final predictions of the front location, represented by
conditional EPs in Fig.~\ref{fig:res1} and Fig.~\ref{fig:res2} as
dashed lines, correspond with one another. In both surveys they yield
a picture of the front being to the west in the southern portions of
the region and gradually bending off toward the north east. The amount
of exploration done by Survey 1 which turned north is greater than
Survey 2 which was coming close to the survey area borders in the
south-western corner.

@@

The case study considers the upper water column in the river plume,
represented by a two dimensional grid. Extensions to three-dimensional
domains are not methodologically different, but likely approximate
calculations by concentrating numerical integration on terms in the
vicinity of the autonomous vehicle \citep{fossum18b}. While we did not
consider any temporal effects, which would be relevant on a larger
time scale, we do consider the extension to spatio-temporal modeling
and envision that advection-diffusion equations could be useful
\citep{sigrist2015stochastic,richardson2017sparsity}. The sampling capabilities are restricted by the sensors onboard the AUV, the computational load and navigation resources. The sampling of temperature and salinity variables can be much higher than the grid resolution, but, due to computational and navigation resources
available on the AUV, assimilation into a discretized fixed grid is necessary.For more complex
oceanographic phenomena, the methods will need to be extended to
non-Gaussian phenomena, possibly feature-based mixtures of Gaussian
processes which could potentially be run onboard augmented by
dynamical models. Running numerical models onboard a robotic vehicle
is currently infeasible, but high-resolution ocean models or remote
sensing data can be used to fit a more complex statistical model
\cite{davidson19}.

>>>>

The final predictions of the front location, represented by
conditional EPs in Fig.~\ref{fig:res1} and Fig.~\ref{fig:res2} as
dashed lines, correspond with one another. In both surveys they yield
a picture of the front being to the west in the southern portions of
the region and gradually bending off toward the north east. The amount
of exploration done by Survey 1 which turned north is greater than
Survey 2 which was coming close to the survey area borders in the
south-western corner.
The AUV paths in both runs are relatively winding in the sense that the vehicle reacts much to the data. In our experience from simulated runs, the paths tend to be smoother when we use a larger nugget effect. But here the estimated nugget from the actual trial data is close to the specified level.

@@

The case study considers the upper water column in the river plume,
represented by a two dimensional grid. Extensions to three-dimensional
domains are not methodologically different, but likely approximate
calculations by concentrating numerical integration on terms in the
vicinity of the autonomous vehicle \citep{fossum18b}. While we did not
consider any temporal effects, which would be relevant on a larger
time scale, we do consider the extension to spatio-temporal modeling
and envision that advection-diffusion equations could be useful
\citep{sigrist2015stochastic,richardson2017sparsity}. The sampling capabilities are restricted by the sensors onboard the AUV, the computational load, and navigation resources. The sampling of temperature and salinity variables can be much higher than the grid resolution, but, due to computational and navigation resources
available on the AUV, assimilation into a discretized fixed grid is necessary. For more complex
oceanographic phenomena, the methods will need to be extended to
non-Gaussian phenomena, possibly feature-based mixtures of Gaussian
processes which could potentially be run onboard augmented by
dynamical models. Running numerical models onboard a robotic vehicle
is currently infeasible, but high-resolution ocean models or remote
sensing data can be used to fit a more complex statistical model
\cite{davidson19}.
